Weekday sleep, total cholestrol and the modifying effect of age on systolic blood pressure: An analysis of the 2015-2018 non-obese NHANES population

### 2023-12-11 manuscript/paper.Rmd
hi

Univariate regression
analysis results exploring the association between the exposures,
selected potential confounding variables showed substantial associations between BMI ($\beta = 0.95$ mmHg; 95%CI [0.60 - 1.3]; p<0.05), TC ($\beta = 3.6$ mmHg; 95%CI [2.9 - 4.3]; p<0.05), albumin ($\beta = -0.48$ mmHg; 95% CI [-0.68mmHg - -0.27]; p < 0.05), AST ($\beta = 0.12$ mmHg; 95%CI [0.07 - 0.18]; p <0.05), and age (years) ($\beta = 0.49$ mmHg; 95% CI [0.45 - 0.52]; p <0.05) and systolic blood pressure. Weekday sleep duration between 6 and 8 hours was associated with a 4.0 mmHg (95% CI [-6.8 - -1.2]) decrease in the average systolic blood pressure compared to 6 hours of sleep. On the other hand, more than 8 hours of sleep was associated with a decrease of 3.1 mmHg in systolic blood compared to 6 hours of sleep. Substantial associations were noted between citizenship, education, number of children below five years, gender, marital status, race, smoking status, snorting, alcohol, and income.

@@

Table \@ref(tab:multreg) shows the coefficients of sleep, age, total
cholesterol and the interaction between total cholesterol and age

@@

\begin{table}[H]
\caption{\label{tab:multreg} Sleep, Age and Cholesterol Coefficients for Model 5}
\centering
\begin{tabular}[t]{lrrr}
\toprule
  \textbf{Variable} & \textbf{Estimate} & \textbf{95\% CI}  & \textbf{p-value} \\
\midrule
\textbf{Intercept} & 102.55 &(88.42,116.69) & $<0.001$\\
\textbf{Sleep} & 0.75 & (0.50, 1.00) & 0.002 \\
$<6$hrs &  &  &  \\
$>8$hrs & -1.4 & (-3.7,0.92) & 0.2\\
$6-8$hrs & -3.0 & (-5.3,-0.67) &  0.012\\
\textbf{Age} & -0.04& (-0.22, 0.13)	& 0.6 \\
\textbf{TC} & -3.5	& (-5.2, -1.8) & <0.001 \\
\textbf{TC*Age} & 0.11	& (0.07, 0.14) & <0.001 \\
\bottomrule
\end{tabular}
\end{table}
```

```{r tabfour, echo=FALSE, eval=FALSE}
## Final selected multiple linear regression model
tbl_regression(adj.sbp.int2,
               label = list(bmi ~ "BMI",
                            #hdl ~ "HDL",
                            total_chol ~ "TC",
                            #hemoglobin ~ "Hemoglobin",
                            albumin ~ "Albumin",
                            creatinine ~ "Creatinine",
                            #hypertension ~ "Hypertension",
                            #diabetes ~ "Diabetes",
                            citizenship_status ~ "Citizenship",
                            educ_level ~ "Education",
                            children..5 ~ "Children > 5 yrs",
                            age_yr ~ "Age (yrs)",
                            marital_status ~ "Marital status",
                            #cycle ~ "Survey cycle",
                            sleep ~ "Sleep",
                            race ~ "Race",
                            smoke ~ "Smoking status",
                            snort ~ "Snort",
                            alcohol ~ "Alcohol",
                            income_category ~ "Income",
                            gender ~ "Gender"
               )) %>% modify_header(label ~ "**Variable**") %>%
  bold_labels() %>% 
  as_kable_extra(format = "latex", digits = 2, caption = "MODEL 5: Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP). BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)

>>>>

Univariate regression analysis results exploring the association between
the exposures, selected potential confounding variables showed
substantial associations between BMI ($\beta = 0.95$ mmHg; 95%CI [0.60 -
1.3]; p\<0.05), TC ($\beta = 3.6$ mmHg; 95%CI [2.9 - 4.3]; p\<0.05),
albumin ($\beta = -0.48$ mmHg; 95% CI [-0.68mmHg - -0.27]; p \< 0.05),
AST ($\beta = 0.12$ mmHg; 95%CI [0.07 - 0.18]; p \<0.05), and age
(years) ($\beta = 0.49$ mmHg; 95% CI [0.45 - 0.52]; p \<0.05) and
systolic blood pressure. Weekday sleep duration between 6 and 8 hours
was associated with a 4.0 mmHg (95% CI [-6.8 - -1.2]) decrease in the
average systolic blood pressure compared to 6 hours of sleep. On the
other hand, more than 8 hours of sleep was associated with a decrease of
3.1 mmHg in systolic blood compared to 6 hours of sleep. Substantial
associations were noted between citizenship, education, number of
children below five years, gender, marital status, race, smoking status,
snorting, alcohol, and income.

@@

Table 1 displays the estimated coefficients derived from a multiple
regression analysis involving variables. We only show sleep during
weekdays, age, total cholesterol, and the interaction effect between
total cholesterol and age. Notably, the findings reveal that an increase
of one year in age is associated with a marginal decrease in systolic
blood pressure (SBP) by an amount of 0.04mmHg. However, this observed
reduction is deemed statistically insignificant within the confines of
this model. Conversely, an increase of one unit in total cholesterol is
linked to a more pronounced reduction in SBP by a factor of 3.5mmHg,
which attains statistical significance. Furthermore, the interaction
between age and total cholesterol yields a statistically significant
coefficient estimate of 0.11. This interaction effect suggests that as
both age and total cholesterol concurrently increase by one unit, there
is an associated elevation in SBP by a factor of 0.11mmHg. Additionally,
the coefficients attributed to sleep duration bear analogous
interpretations to those observed in the univariate scenario.
Specifically, weekday sleep durations ranging between 6 and 8 hours
demonstrate an average decrease of 3.0mmHg in systolic blood pressure in
comparison to the reference duration of 6 hours of sleep. Conversely,
durations exceeding 8 hours exhibit a decrease of 1.4mmHg in systolic
blood pressure as compared to the 6-hour reference duration.

@@

\begin{table}[H]
\caption{\label{tab:multreg} Sleep, Age and Cholesterol Coefficients for Model 5}
\centering
\begin{tabular}[t]{lrrr}
\toprule
  \textbf{Variable} & \textbf{Estimate} & \textbf{95\% CI}  & \textbf{p-value} \\
\midrule
\textbf{Intercept} & 102.55 &(88.42,116.69) & $<0.001$\\
\textbf{Sleep} & 0.75 & (0.50, 1.00) & 0.002 \\
$<6$hrs &  &  &  \\
$>8$hrs & -1.4 & (-3.7,0.92) & 0.2\\
$6-8$hrs & -3.0 & (-5.3,-0.67) &  0.012\\
\textbf{Age} & -0.04& (-0.22, 0.13) & 0.6 \\
\textbf{TC} & -3.5  & (-5.2, -1.8) & <0.001 \\
\textbf{TC*Age} & 0.11  & (0.07, 0.14) & <0.001 \\
\bottomrule
\end{tabular}
\end{table}
```
```{r tabfour, echo=FALSE, eval=FALSE}
## Final selected multiple linear regression model
tbl_regression(adj.sbp.int2,
               label = list(bmi ~ "BMI",
                            #hdl ~ "HDL",
                            total_chol ~ "TC",
                            #hemoglobin ~ "Hemoglobin",
                            albumin ~ "Albumin",
                            creatinine ~ "Creatinine",
                            #hypertension ~ "Hypertension",
                            #diabetes ~ "Diabetes",
                            citizenship_status ~ "Citizenship",
                            educ_level ~ "Education",
                            children..5 ~ "Children > 5 yrs",
                            age_yr ~ "Age (yrs)",
                            marital_status ~ "Marital status",
                            #cycle ~ "Survey cycle",
                            sleep ~ "Sleep",
                            race ~ "Race",
                            smoke ~ "Smoking status",
                            snort ~ "Snort",
                            alcohol ~ "Alcohol",
                            income_category ~ "Income",
                            gender ~ "Gender"
               )) %>% modify_header(label ~ "**Variable**") %>%
  bold_labels() %>% 
  as_kable_extra(format = "latex", digits = 2, caption = "MODEL 5: Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP). BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)
